The effects of alignment error and alignment filtering on the \sw detection of positive selection

### 2011-07-18 projects/slrsim/paper/slrsim.tex
paper formatting: remove MBE style in favor of double-spaced ms format.

\documentclass{mbe}
\usepackage{natbib,graphicx,dblfloatfix,multirow,array,threeparttable}
\usepackage{xspace}
\setlength{\bibsep}{0.0pt}

@@

\author{{Gregory Jordan and Nick Goldman}\\
\affiliation{European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom}}
\authorlist{Jordan and Goldman}
\shorttitle{Effect of Alignment Error on Detecting Sitewise Selection}

\mbeabstract{%

@@

}

\keywords{Alignment, positive selection, alignment filtering, indels}

\email{greg@ebi.ac.uk}

\mbestyle

@@

Large-scale comparative genomics has been widely employed to identify
shared and unique biological features between organisms of interest to
the biomedical sciences, with applications ranging from evolutionary
annotation transfer \citep{Bork1998Predicting} to phylogenetics
\citep{Murphy2007Using,Prasad2008Confirming} and population genetic
inference \citep{Hurst2009Genetics}. The decreasing cost of DNA
sequencing has triggered a striking increase in the number of model
and non-model organisms with planned genome sequencing projects,
suggesting that the range and scale of comparative genomics
applications will continue to expand
\citep{Green20072x,2007Identification}.

The study of protein evolutionary rates and selective pressures in
particular has flourished as a result of the growth in comparative
genomics datasets. Clusters of closely-related genome sequences across
a wide taxonomic range have led to a better understanding of which
aspects of molecular evolution are variable and which are constant
\citep{Wolf2009Nonlinear}, and an increased sampling of species should
continue to boost the power and accuracy of individual analyses within
a given clade. This is especially beneficial for the calculation of
spatially precise evolutionary estimates, as additional species
sampling has been shown to be an effective means of boosting the
accuracy and power of \sw detection of positive selection and
evolutionary constraint
\citep{Anisimova2001Accuracy,Massingham2005Detecting}. Site-specific
evolutionary estimates have proved especially valuable when analyzed
in conjunction with other protein-based datasets such as protein
structural features \citep{Lin2007Proportion,Ramsey2011Relationship},
human population diversity \citep{2010Map}, and human disease
mutations \citep{Arbiza2006Selective}.

A major concern in the \sw detection of protein positive
selection is that the effect of alignment error is not well
characterized. Intuitively, one might expect alignment error to result
mainly in an increased number of false positives, as the spurious
alignment of nonhomologous codons on average would result in a high
number of apparent non-synonymous substitutions and a low number of
synonymous substitutions (since a randomly chosen pair of codons are
more likely to be nonsynonymous than synonymous). However, false
negatives may also be introduced, either through over-alignment of
synonymous codons (increasing the synonymous substitution rate) or
non-alignment of homologous codons (causing reduced power due to less
evolutionary information). Since different aligners employ a variety
of algorithms, evolutionary models, and heuristic optimizations
\citep{Notredame2007Recent}, each program may be more or less prone to different
types of alignment error, causing potentially large variations in the
nature and magnitude of its impact on the detection of positive
selection.

@@

The protein structure and the evolutionary divergence of a dataset may
also contribute to the effects of alignment error. Differently
structured protein regions show variable tolerance to biological
indels, with indels more common in extracellular and transmembrane
proteins than in highly folded enzymes and housekeeping genes
\citep{delaChaux2007DNA} . This suggests that well-folded protein
regions will experience fewer biological indels---and will therefore
be less susceptible to alignment error---than less structured
regions. Furthermore, the evolutionary divergence of a dataset affects
the power of \sw inference and the prevalence of indels in different
ways. Because maximum-likelihood methods for detecting positive
selection require data in the form of observed substitution events,
they show little power at low divergence and their highest power at
intermediate to high divergence levels
\citep{Anisimova2001Accuracy}. On the other hand, the number of indels
in a dataset should increase with divergence. These two trends suggest
that the overall power will be low at both extremes of divergence,
with little inference power at low divergence (due to the scarcity of
data in the form of observed substitutions) and an overwhelming amount
of alignment error at high divergence (due to the large number of
indel events). It is important to note, however, that the majority of
genes in many biological clades of interest (such as mammals,
vertebrates, fruit flies, and yeast) fall within the middle range of
divergences where \sw methods are at their most powerful and where
multiple alignment is a difficult---but not hopeless---problem. As
such, it is important to seek an understanding of the impact of
alignment error on overall error rates within this important range of
divergence levels.

@@

A limitation to the analysis of error in empirical datasets is the
lack of a benchmark set of true alignments and positively-selected
sites. Markova-Raina and Petrov \citeyearpar{Markova-Raina2011High}
used the expected general effect of alignment error (an increase in
false positives due to misalignment of non-homologous codons) as a
proxy by which to compare different methods, allowing for the
conclusion that PRANK was the least error-prone aligner in their
analysis. However, the absolute number of false positives remained
uncertain due to the possible conflation of multiple sources of error:
in addition to alignment error, the authors noted that gene
mis-annotation was responsible for many apparent false positives, and
there is also an expected error rate from the likelihood inference
method itself. This limitation leaves important and interesting
questions regarding the nature of alignment error and its quantitative
impact on the detection of positive selection left unanswered.

Controlled simulation experiments provide a natural framework for
investigating error rates in detail, allowing one to pinpoint the
source of errors in multi-step analyses such as alignment followed by
evolutionary inference. This approach has widely been employed in
assessing the robustness of phylogenetic inference methods to
misalignment
\citep{Dwivedi2009Phylogenetic,Ogden2006Multiple,Loytynoja2008PhylogenyAware}
but those results cannot be easily extrapolated to the analysis of \sw
selective pressures. More recently, Fletcher and Yang
\citeyearpar{Fletcher2010Effect} performed a series of simulation
experiments investigating alignment error in the use of branch-site
models to detect positive selection in genes. Their results showed
that most aligners caused false positives by over-aligning codons and
that datasets from mammalian and vertebrate gene families contain
enough evolutionary divergence to make false positive errors resulting
from misalignment a legitimate concern.

Reflecting a widespread awareness of the problem of misalignment,
methods for identifying and removing uncertain or unreliable alignment
regions have been commonly used in phylogenetic and molecular
evolutionary analyses. The popular Gblocks applies a set of heuristic
criteria to identify conserved blocks deemed suitable for phylogenetic
or evolutionary analysis \citep{Castresana2000Selection} while a
number of aligners such as T-Coffee \citep{Notredame2000TCoffee} and
PRANK \citep{Loytynoja2005From} produce estimates of alignment
confidence or reliability. Alignment confidence scores based on the
robustness of alignments to guide tree perturbations
\citep{Penn2010Alignment} or the sampling of suboptimal alignments
\citep{Kim2011PSAR} have also been proposed. Unfortunately, despite
their widespread use, the impact of the many available filtering
methods on phylogenetic and evolutionary analyses has not been well
studied. Even for a single filtering program, Gblocks, results have
been contradictory: one simulation-based study found that it improved
the phylogenetic signal \citep{Talavera2007Improvement} while an
empirical study across a wide range of taxa found that
Gblocks-filtered alignments produced worse phylogenetic trees than
unfiltered alignments \citep{Dessimoz2010Phylogenetic}. With the
application of published filtering methods to alignments before
testing for positive selection becoming standard practice
\citep{Studer2008Pervasive,Aguileta2009Rapidly}, an analysis of the
benefit---if any---of alignment filtering to the detection of positive
selection would seem well-warranted.

@@

An overview of the simulation parameters used in this study can be
found in Table \ref{table_1}. Three rooted trees were used to guide
the simulation of protein-coding DNA alignments: the artificial
6-taxon tree used by Anisimova and Yang
\citeyearpar{Anisimova2001Accuracy} and Massingham and Goldman
\citeyearpar{Massingham2005Detecting} rooted at its midpoint, the
17-taxon vertebrate B-globin tree from Yang et
al. \citeyearpar{Yang2000CodonSubstitution}, and the 44-taxon
vertebrate tree used by the ENCODE project
\citep{2007Identification,Nikolaev2007Early}. Trees, shown with their
original branch lengths in Figure \ref{fig_1}, were scaled to
comparable divergence levels by normalizing their mean path length
(MPL), defined as the root-to-tip branch length averaged across all
lineages in the tree. We simulated alignments with MPL divergence
between 0.05 and 2.0 synonymous substitutions per synonymous site,
spanning the range of evolutionary divergences observed in several
clades of organisms with fully-sequenced genomes (Table
\ref{table_2}).
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.5]{fig1.pdf}
\end{center}
\caption{Phylogenetic trees used for simulation and analysis. The
  original scale for each tree is indicated by a scale bar, but trees
  were scaled to equal mean path length (MPL) divergence levels for
  simulation. (a) A 6-taxon artificial tree used in previous
  simulations
  \citep{Anisimova2001Accuracy,Massingham2005Detecting}. (b) A tree
  estimated from $\beta$-globin genes of 17 vertebrates and used in
  previous empirical analyses and simulation studies
  \citep{Anisimova2001Accuracy,Anisimova2002Accuracy}. (c) The
  44-species tree used by the ENCODE project
  \citep{2007Identification,Nikolaev2007Early}. The nodes indicated by
  arrows were used as the reference species when comparing the true
  and inferred alignment (see Methods).}
\label{fig_1}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.75]{fig2.pdf}
\end{center}
\caption{Alignments were simulated without indels for three tree
  shapes and analyzed with SLR, PAML M8, or PAML M2. Fifty replicate
  alignments were simulated for each data point. The performance of
  each analysis method, as measured by four summary statistics, is
  plotted as a function of the mean path length (MPL) divergence. From
  top to bottom: true positive rate (TPR) at the recommended cutoff
  threshold (0.95 for PAML and 3.84 for SLR); false positive rate
  (FPR) at the recommended cutoff threshold; TPR at a 1\% FPR
  threshold; Pearson's correlation coefficient between the true and
  inferred sitewise $\omg$.}
\label{fig_2}
\end{figure}

@@

When the indel rate was greater than zero, performance levels varied
significantly for different tree sizes, alignment algorithms, and
evolutionary divergences. Figure \ref{fig_3} shows the same
performance measurements as Figure \ref{fig_2} for simulations without
indels (gray lines, Figure \ref{fig_2}) and with indels (black lines,
Figure \ref{fig_2}) analyzed using three different aligners and the
true alignment. For the indel simulations, the indel rate was held
constant at 0.1 indel event per substitution.
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.75]{fig3.pdf}
\end{center}
\caption{Sequences were simulated without indels (gray lines) or with
  indels (black lines) using one of three tree shapes, aligned with
  one of three aligners, and analyzed with SLR; true alignments were
  separately analyzed with SLR (solid black lines). One hundred
  replicate alignments were simulated for each data point. The
  performance of each dataset, as measured by four summary statistics,
  is plotted as a function of the mean path length (MPL)
  divergence. From top to bottom: true positive rate (TPR) at the
  recommended cutoff threshold; false positive rate (FPR) at the
  recommended cutoff threshold; TPR at a 1\% FPR threshold; Pearson's
  correlation coefficient between the true and inferred $\omg$.}
\label{fig_3}
\end{figure}

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[scale=0.8]{fig4.pdf}
\end{center}
\caption{Sequences were simulated with indels using one of three tree
  shapes and a range of indel rates and mean path length (MPL)
  divergence levels. Alignments were inferred with one of four
  aligners and analyzed with SLR; true alignments were separately
  analyzed with SLR. One hundred replicates were simulated for each
  set of conditions. Each square is colored according to the
  performance at a given (indel rate, MPL) pair as measured by one of
  three summary statistics: (a) the true positive rate (TPR) at the
  recommended cutoff threshold, (b) the false positive rate (FPR) at
  the recommended cutoff threshold, or (c) the TPR at a 1\% FPR
  threshold.}
\label{fig_4}
\end{figure*}

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[scale=0.65]{fig5.pdf}
\end{center}
\caption{Sequences were simulated using the 17-taxon tree and a range
  of indel rates and mean path length (MPL) divergence
  levels. Alignments were inferred using (a) ClustalW or (b) \prankc,
  either left unfiltered or filtered with one of four alignment
  filters, and analyzed with SLR; true alignments were left unfiltered
  and separately analyzed with SLR. One hundred fifty replicates were
  simulated for each set of conditions. Squares are colored according
  to the ratio of the performance of the indicated filter to the
  performance of the unfiltered ClustalW or \prankc alignment as
  measured by one of four summary statistics. In columns from left to
  right: the number of unfiltered (i.e., non-N) residues remaining
  in the alignment; the false positive rate (FPR) at the recommended
  cutoff threshold; the true positive rate (TPR) at the recommended
  cutoff threshold; the TPR at a 1\% FPR threshold (\tpr). Note that the
  maximum percentage of filtered residues was capped at 50\% for all
  methods except Gblocks.}
\label{fig_5}
\end{figure*}

>>>>

\documentclass{article}
\usepackage{natbib,graphicx,dblfloatfix,multirow,array,threeparttable}
\usepackage{xspace,bm,amsmath,amssymb}
\usepackage[nolists,tablesfirst,nomarkers]{endfloat}

\usepackage[a4paper]{geometry}

\linespread{1.5}

\setlength{\bibsep}{0.0pt}

@@

\author{{Gregory Jordan and Nick Goldman}\\
\affiliation{European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom}}
%\authorlist{Jordan and Goldman}
%\shorttitle{Effect of Alignment Error on Detecting Sitewise Selection}

%\mbeabstract{% 
\section*{Abstract}

@@

%}

%\keywords{Alignment, positive selection, alignment filtering, indels}

%\email{greg@ebi.ac.uk}

%\mbestyle

@@

The decreasing cost of DNA sequencing has triggered a striking
increase in the number of model and non-model organisms with planned
genome sequencing projects, suggesting that the range and scale of
comparative genomics applications will continue to expand
\citep{Green20072x,2007Identification}. These clusters of
closely-related genome sequences across a wide taxonomic range have
led to a better understanding of which aspects of molecular evolution
are variable and which are constant \citep{Wolf2009Nonlinear} and an
increased sampling of species should continue to boost the power and
accuracy of individual analyses within a given clade.

The study of protein evolutionary rates and selective pressures in
particular has flourished as a result of the growth in comparative
genomics datasets. This is especially beneficial for the calculation
of spatially precise evolutionary estimates, as additional species
sampling has been shown to be an effective means of boosting the
accuracy and power of \sw detection of positive selection and
evolutionary constraint
\citep{Anisimova2001Accuracy,Massingham2005Detecting}. Site-specific
evolutionary estimates have proved especially valuable when analyzed
in conjunction with other protein-based datasets such as protein
structural features \citep{Lin2007Proportion,Ramsey2011Relationship},
human population diversity \citep{2010Map} and human disease
mutations \citep{Arbiza2006Selective}.

A major concern in the detection of protein positive selection is that
the effect of alignment error is not well characterized. Intuitively,
one might expect alignment error to result mainly in an increased
number of false positives, as the spurious alignment of nonhomologous
codons on average would result in a high number of apparent
non-synonymous substitutions and a low number of synonymous
substitutions (since a randomly chosen pair of codons are more likely
to be nonsynonymous than synonymous). However, false negatives may
also be introduced, either through over-alignment of synonymous codons
(increasing the synonymous substitution rate) or non-alignment of
homologous codons (causing reduced power due to less evolutionary
information). Since different aligners employ a variety of algorithms,
evolutionary models, and heuristic optimizations
\citep{Notredame2007Recent}, each program may be more or less prone to
different types of alignment error, causing potentially large
variations in the nature and magnitude of its impact on the detection
of positive selection.

@@

The protein structure and the evolutionary divergence of a dataset may
also contribute to the effects of alignment error. Differently
structured protein regions show variable tolerance to biological
indels, with indels more common in extracellular and transmembrane
proteins than in highly folded enzymes and housekeeping genes
\citep{delaChaux2007DNA} . This suggests that well-folded protein
regions will experience fewer biological indels---and will therefore
be less susceptible to alignment error---than less structured
regions.

Furthermore, the evolutionary divergence of a dataset affects the
power of \sw inference and the prevalence of indels in different
ways. As maximum-likelihood methods for detecting positive selection
require data in the form of observed substitution events, they show
little power at low divergence and their highest power at intermediate
to high divergence levels \citep{Anisimova2001Accuracy}. However,
alignment error should be greatest at high divergences, which may have
the effect of reducing power. These two trends suggest that the
overall power will be low at both extremes of divergence, with little
inference power at low divergence (due to the scarcity of data in the
form of observed substitutions) and an overwhelming amount of
alignment error at high divergence (due to the large number of indel
events).

Luckily the majority of genes in many biological clades of interest
(such as mammals, vertebrates, fruit flies, and yeast) fall within the
middle range of divergences where \sw methods are at their most
powerful and where multiple alignment is a difficult---but not
hopeless---problem. As such, it is important to seek an understanding
of the impact of alignment error on overall error rates within this
important range of divergence levels.

@@

A limitation to the analysis of error in empirical datasets is the
lack of a benchmark set of true alignments and positively-selected
sites. Markova-Raina and Petrov \citeyearpar{Markova-Raina2011High}
used their expected general effect of alignment error (an increase in
false positives due to misalignment of non-homologous codons) as a
proxy by which to compare different methods, allowing for the
conclusion that PRANK was the least error-prone aligner in their
analysis. However, the absolute number of false positives remained
uncertain and there was the possibility of conflating multiple sources
of error: in addition to alignment error, the authors noted that gene
mis-annotation was responsible for many apparent false positives, and
there is also an expected error rate from the likelihood inference
method itself. This limitation leaves important and interesting
questions regarding the nature of alignment error and its quantitative
impact on the detection of positive selection left unanswered by
empirical studies.

Controlled simulation experiments provide a natural framework for
investigating error rates in detail, allowing one to pinpoint the
sources of error in multi-step analyses such as alignment followed by
evolutionary inference. This approach has been employed in assessing
the robustness of phylogenetic inference methods to misalignment
\citep{Dwivedi2009Phylogenetic,Ogden2006Multiple,Loytynoja2008PhylogenyAware}
but those results cannot be easily extrapolated to the analysis of \sw
selective pressures. More recently, Fletcher and Yang
\citeyearpar{Fletcher2010Effect} performed a series of simulation
experiments investigating alignment error in the use of the
branch-site test to detect positive selection in genes. Their results
showed that most aligners caused false positives by over-aligning
codons and that datasets from mammalian and vertebrate gene families
contain enough evolutionary divergence to make false positive errors
resulting from misalignment a legitimate concern.

Reflecting a widespread awareness of the problem of misalignment,
methods for identifying and removing uncertain or unreliable alignment
regions have been commonly used in phylogenetic and molecular
evolutionary analyses. The popular Gblocks program applies a set of
heuristic criteria to identify conserved blocks deemed suitable for
phylogenetic or evolutionary analysis \citep{Castresana2000Selection}
while a number of aligners such as T-Coffee
\citep{Notredame2000TCoffee} and PRANK \citep{Loytynoja2005From}
produce estimates of alignment confidence or reliability. Alignment
confidence scores based on the robustness of alignments to guide tree
perturbations \citep{Penn2010Alignment} or the sampling of suboptimal
alignments \citep{Kim2011PSAR} have also been proposed. Unfortunately,
despite their widespread use, the impact of the many available
filtering methods on phylogenetic and evolutionary analyses has not
been well studied. Even for a single filtering program, Gblocks,
results have been contradictory: one simulation-based study found that
it improved the phylogenetic signal \citep{Talavera2007Improvement}
while an empirical study across a wide range of taxa found that
Gblocks-filtered alignments produced worse phylogenetic trees than
unfiltered alignments \citep{Dessimoz2010Phylogenetic}. With the
application of published filtering methods to alignments before
testing for positive selection becoming standard practice
\citep{Studer2008Pervasive,Aguileta2009Rapidly}, an analysis of the
benefits of alignment filtering to the detection of positive selection
would seem well-warranted.

@@

An overview of the simulation parameters used in this study can be
found in Table \ref{table_1}. Three rooted trees were used to guide
the simulation of protein-coding DNA alignments: the artificial
6-taxon tree used by Anisimova and Yang
\citeyearpar{Anisimova2001Accuracy} and Massingham and Goldman
\citeyearpar{Massingham2005Detecting} rooted at its midpoint, the
17-taxon vertebrate B-globin tree from Yang et
al. \citeyearpar{Yang2000CodonSubstitution}, and the 44-taxon
vertebrate tree used by the ENCODE project
\citep{2007Identification,Nikolaev2007Early}. Trees, shown with their
original branch lengths in Figure \ref{fig_1}, were scaled to
comparable divergence levels by normalizing their mean path length
(MPL), defined as the root-to-tip branch length averaged across all
lineages in the tree. We simulated alignments with MPL divergence
between 0.05 and 2.0 synonymous substitutions per synonymous site,
spanning the range of evolutionary divergences observed in several
clades of organisms with fully-sequenced genomes (Table
\ref{table_2}).
\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.5]{fig1.pdf}
\includegraphics[scale=0.8]{fig1.pdf}
\end{center}
\caption{Phylogenetic trees used for simulation and analysis. The
  original scale for each tree is indicated by a scale bar, but trees
  were scaled to equal mean path length (MPL) divergence levels for
  simulation. (a) A 6-taxon artificial tree used in previous
  simulations
  \citep{Anisimova2001Accuracy,Massingham2005Detecting}. (b) A tree
  estimated from $\beta$-globin genes of 17 vertebrates and used in
  previous empirical analyses and simulation studies
  \citep{Anisimova2001Accuracy,Anisimova2002Accuracy}. (c) The
  44-species tree used by the ENCODE project
  \citep{2007Identification,Nikolaev2007Early}. The nodes indicated by
  arrows were used as the reference species when comparing the true
  and inferred alignment (see Methods).}
\label{fig_1}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.75]{fig2.pdf}
\includegraphics[scale=1.25]{fig2.pdf}
\end{center}
\caption{Alignments were simulated without indels for three tree
  shapes and analyzed with SLR, PAML M8, or PAML M2. Fifty replicate
  alignments were simulated for each data point. The performance of
  each analysis method, as measured by four summary statistics, is
  plotted as a function of the mean path length (MPL) divergence. From
  top to bottom: true positive rate (TPR) at the recommended cutoff
  threshold (0.95 for PAML and 3.84 for SLR); false positive rate
  (FPR) at the recommended cutoff threshold; TPR at a 1\% FPR
  threshold; Pearson's correlation coefficient between the true and
  inferred sitewise $\omg$.}
\label{fig_2}
\end{figure}

@@

When the indel rate was greater than zero, performance levels varied
significantly for different tree sizes, alignment algorithms, and
evolutionary divergences. Figure \ref{fig_3} shows the same
performance measurements as Figure \ref{fig_2} for simulations without
indels (gray lines, Figure \ref{fig_2}) and with indels (black lines,
Figure \ref{fig_2}) analyzed using three different aligners and the
true alignment. For the indel simulations, the indel rate was held
constant at 0.1 indel event per substitution.
\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.75]{fig3.pdf}
\includegraphics[scale=1.25]{fig3.pdf}
\end{center}
\caption{Sequences were simulated without indels (gray lines) or with
  indels (black lines) using one of three tree shapes, aligned with
  one of three aligners, and analyzed with SLR; true alignments were
  separately analyzed with SLR (solid black lines). One hundred
  replicate alignments were simulated for each data point. The
  performance of each dataset, as measured by four summary statistics,
  is plotted as a function of the mean path length (MPL)
  divergence. From top to bottom: true positive rate (TPR) at the
  recommended cutoff threshold; false positive rate (FPR) at the
  recommended cutoff threshold; TPR at a 1\% FPR threshold; Pearson's
  correlation coefficient between the true and inferred $\omg$.}
\label{fig_3}
\end{figure}

@@

\begin{figure*}[t]
\begin{center}
%\includegraphics[scale=0.8]{fig4.pdf}
\includegraphics[scale=0.8]{fig4.pdf}
\end{center}
\caption{Sequences were simulated with indels using one of three tree
  shapes and a range of indel rates and mean path length (MPL)
  divergence levels. Alignments were inferred with one of four
  aligners and analyzed with SLR; true alignments were separately
  analyzed with SLR. One hundred replicates were simulated for each
  set of conditions. Each square is colored according to the
  performance at a given (indel rate, MPL) pair as measured by one of
  three summary statistics: (a) the true positive rate (TPR) at the
  recommended cutoff threshold, (b) the false positive rate (FPR) at
  the recommended cutoff threshold, or (c) the TPR at a 1\% FPR
  threshold.}
\label{fig_4}
\end{figure*}

@@

\begin{figure*}[t]
\begin{center}
%\includegraphics[scale=0.65]{fig5.pdf}
\includegraphics[scale=0.65]{fig5.pdf}
\end{center}
\caption{Sequences were simulated using the 17-taxon tree and a range
  of indel rates and mean path length (MPL) divergence
  levels. Alignments were inferred using (a) ClustalW or (b) \prankc,
  either left unfiltered or filtered with one of four alignment
  filters, and analyzed with SLR; true alignments were left unfiltered
  and separately analyzed with SLR. One hundred fifty replicates were
  simulated for each set of conditions. Squares are colored according
  to the ratio of the performance of the indicated filter to the
  performance of the unfiltered ClustalW or \prankc alignment as
  measured by one of four summary statistics. In columns from left to
  right: the number of unfiltered (i.e., non-N) residues remaining
  in the alignment; the false positive rate (FPR) at the recommended
  cutoff threshold; the true positive rate (TPR) at the recommended
  cutoff threshold; the TPR at a 1\% FPR threshold (\tpr). Note that the
  maximum percentage of filtered residues was capped at 50\% for all
  methods except Gblocks.}
\label{fig_5}
\end{figure*}
